---
title: "Read and plot GLODAPv2 data"
author: "Donghe Zhu and Dr Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options}
# Some basic settings to improve the html output
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


# Required Packages

```{r Load_package}
library(tidyverse) # the tidverse combines several packages
library(ncdf4)
library(stars)
library(tidync)
library(tidyselect)
library(geosphere)
library(PCICt)
library(lubridate)
```


```{r ggplot_theme}
# ggplot theme for all plots
theme_set(theme_bw())
```

# The problem I encounter here is that I want to extract the month from model data. First, I convert from Pcict to Date form, but the process takes hours and I don't know why.
# Read and preprocess data from netcdf file
```{r read_file_and_extract_data, eval=FALSE}

# read DIC from netcdf file
nc_data <- read_ncdf(here::here("data",
                                "GIAF_JRA.pop.h.0354.nc"),
                     var = "DIC")

# convert to tibble
nc_data_tibble <- nc_data %>%
  as.tibble() %>%
  mutate(nlon = if_else(nlon < 20, nlon + 360, nlon)) %>%
  mutate(z_t = z_t / 100) %>%
  rename(lon = nlon, lat = nlat, depth = z_t) %>%
  mutate(DIC = as.numeric(DIC)) %>%
  filter(!is.na(DIC))

# convert time to month

nc_data_tibble <- nc_data_tibble %>%
  mutate(time = as.Date.character(time))
nc_data_tibble <- nc_data_tibble %>%
  mutate(time = month(time) + 1) %>%
  mutate(time = if_else(time == 13, 1, time))
  rename(month = time)
# or use
nc_data_tibble <- nc_data_tibble %>%
  mutate(time = as.POSIXct(time)) %>%
  mutate(time = month(time) + 1) %>%
  mutate(time = if_else(time == 13, 1, time))
  rename(month = time)
  
nc_data_tibble %>% 
  write_csv(here::here("data",
                       "GIAF_JRA.pop.h.0354_as_date.nc"))

```

# Read data from netcdf and csv file
```{r Read_file_and_extract_data, eval=FALSE}
# read csv file
GLODAP <-
  read_csv(
    here::here(
      "data",
      "GLODAPv2.2020_clean_2007.csv"
    ))

landmask <- 
  read_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                       "land_mask_WOA18.csv"))

# convert time to month
GLODAP_clean <- GLODAP %>%
  mutate(date = month(date)) %>%
  rename(month = date)un

# plot GLODAP sites
ggplot() + 
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_point(data = GLODAP_clean,
             aes(lon, lat)) 

```

# Besides, for subseting the model data with observations based on near lat, lon, depth and month. I wonder if you have some code or references, I find it kind of difficult to make it by myself.


# This part is the problem I encounter in read_model_data.RMD and I just copy here. When I extract all the interested data from the file and filter the NaN value, I always encounter a erre r: Error: vector memory exhausted (limit reached?). I wonder how could I avoid it.

### Read netcdf
```{r read_file, eval=FALSE}
# read one variable from file
nc_data <- read_ncdf(here::here("data",
                                "GIAF_JRA.pop.h.0354.nc"),
                     var = c("DIC", "ALK", "O2", "AOU", "NO3", "PO4", "SALT", "TEMP", "SiO3"))

# convert to tibble
nc_data_tibble <- nc_data %>%
  as.tibble() %>%
  mutate(nlon = if_else(nlon < 20, nlon + 360, nlon)) %>%
  mutate(z_t = z_t / 100) %>%
  rename(lon = nlon, lat = nlat, depth = z_t)

# convert as numeric
nc_data_tibble <- nc_data_tibble %>%
  mutate(DIC = as.numeric(DIC)) %>%
  mutate(ALK = as.numeric(ALK)) %>%
  mutate(O2 = as.numeric(O2)) %>%
  mutate(AOU = as.numeric(AOU)) %>%
  mutate(NO3 = as.numeric(NO3)) %>%
  mutate(PO4 = as.numeric(PO4)) %>%
  mutate(SALT = as.numeric(SALT)) %>%
  mutate(TEMP = as.numeric(TEMP)) %>%
  mutate(SiO3 = as.numeric(SiO3))

# filter NaN values
nc_data_tibble <- nc_data_tibble %>%
  filter(!is.na(DIC))

# filter surface data
nc_data_tibble_surface <- nc_data_tibble %>%
  filter(depth == 5)

# filter August data
nc_data_tibble_surface_Aug <- nc_data_tibble_surface %>%
  filter(time == as.PCICt("0355-07-30", cal = "365_day")) 

```
