---
title: "Read and plot model data"
author: "Donghe Zhu and Dr Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
# Some basic settings to improve the html output
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r library, message=FALSE, warning=FALSE}
library(tidyverse) # the tidverse combines several packages
library(ncdf4)
library(stars)
library(tidync)
library(tidyselect)
library(geosphere)
library(PCICt)
library(patchwork)
```


```{r read_parameters, include = FALSE}
parameters <- data.frame(
  lon_Atl_section = 335.5,
  lat_section_N = 59.5,
  lat_section_S = -59.5,
  lon_Pac_section = 189.5,
  lon_middle = 50.5,
  lat_max = 65,
  lat_min = -85
)
```

```{r ggplot_theme}
# ggplot theme for all plots
theme_set(theme_bw())
```

# Data source
- UP Model results
- [World Ocean Atlas 2018](https://www.nodc.noaa.gov/OC5/woa18/woa18data.html){target="_blank"}

# Masks


## Land

### Read mask

The land sea mask with 1x1° resolution from the file `landsea_01.msk` was used.

```{r read_landsea_mask}

landsea_01 <- read_csv(here::here("data/World_Ocean_Atlas_2018",
                                  "landsea_01.msk"),
                         skip = 1,
                         col_types = list(.default = "d"))

```

### Label

According to the [WOA18 documentation](https://data.nodc.noaa.gov/woa/WOA18/DOC/woa18documentation.pdf){target="_blank"} document:  

*"The landsea_XX.msk contains the standard depth level number at which the bottom of the ocean is first encountered at each quarter-degree or one-degree square for the entire world.  Land will have a value of 1, corresponding to the surface."*

The landmask was derived as coordinates with value 1.

```{r assign_region_labels}

landmask <- landsea_01 %>% 
  mutate(region = if_else(Bottom_Standard_Level == "1",
                         "land", "ocean")) %>% 
  select(-Bottom_Standard_Level)

landmask <- landmask %>% 
  rename(lat = Latitude,
         lon = Longitude) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

landmask <- landmask %>%
  filter(region == "land", 
         lat >= parameters$lat_min,
         lat <= parameters$lat_max) %>% 
  select(-region)

rm(landsea_01)
```

## Basins

### Read mask

The surface mask (0m) with 1x1° resolution from the file `basinmask_01.msk` was used.

```{r read_basinmask}

basinmask_01 <- read_csv(here::here("data/World_Ocean_Atlas_2018",
                                    "basinmask_01.msk"),
                         skip = 1,
                         col_types = list(.default = "d"))

basinmask_01 <- basinmask_01 %>% 
  select(Latitude:Basin_0m) %>% 
  mutate(Basin_0m = as.factor(Basin_0m)) %>% 
  rename(lat = Latitude, lon = Longitude)
  
```


### Labels

According to [WOA FAQ](https://www.nodc.noaa.gov/OC5/WOD/wod-woa-faqs.html){target="_blank"} website and [WOA18 documentation](https://data.nodc.noaa.gov/woa/WOA18/DOC/woa18documentation.pdf){target="_blank"}, number codes in the mask files were used to assign ocean basins as follows:  

Atlantic Ocean:  

- 1: Atlantic Ocean
- 10: Southern Ocean between 63°W and 20°E
- 11: Arctic Ocean (restricted by northern latitude limit `r parameters$lat_max`N)

Indian Ocean:  

- 3: Indian Ocean 
- 10: Southern Ocean between 20°E and 147°E

Pacific Ocean:  

- 2: Pacific Ocean
- 10: Southern Ocean between 147°E and 63°W
- 12: Sea of Japan
- 56: Bay of Bengal

For eMLR model fitting and mapping, Indian and Pacific Ocean were combined as Indo-Pacific.

```{r assign_basin_labels}

basinmask_01 <- basinmask_01 %>% 
  filter(Basin_0m %in% c("1", "2", "3", "10", "11", "12", "56"),
         lat <= parameters$lat_max) %>% 
  mutate(basin_AIP = "none",
         basin_AIP = case_when(
           Basin_0m == "1" ~ "Atlantic",
           Basin_0m == "10" & lon >= -63 & lon < 20 ~ "Atlantic",
           Basin_0m == "11" ~ "Atlantic",
           Basin_0m == "3" ~ "Indian",
           Basin_0m == "10" & lon >= 20 & lon < 147 ~ "Indian",
           Basin_0m == "2" ~ "Pacific",
           Basin_0m == "12" ~ "Pacific",
           Basin_0m == "56" ~ "Pacific",
           Basin_0m == "10" & lon >= 147 | lon < -63 ~ "Pacific")) %>% 
  mutate(basin = if_else(basin_AIP == "Atlantic",
                         "Atlantic",
                         "Indo-Pacific")) %>% 
  select(-Basin_0m) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

### Map

```{r basin_masks_WOA18_map, fig.asp=0.6}

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_raster(data = basinmask_01,
              aes(lon, lat, fill = basin_AIP)) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        legend.title = element_blank(),
        axis.title = element_blank())

```


## Global section

To plot sections from the North Atlantic south to the Southern Ocean, around Antartica and back North across the Pacific Ocean, corresponding coordinates were subsetted from the basin mask and distances between coordinate grid points calculated.

```{r create_section_global}

section <- basinmask_01 %>%
  select(lon, lat)

Atl_NS <- section %>%
  filter(
    lon == parameters$lon_Atl_section,
    lat <= parameters$lat_section_N,
    lat >= parameters$lat_section_S
  ) %>%
  arrange(-lat)

Atl_SO <- section %>%
  filter(lon > parameters$lon_Atl_section,
         lat == parameters$lat_section_S) %>%
  arrange(lon)

Pac_SO <- section %>%
  filter(lon < parameters$lon_Pac_section,
         lat == parameters$lat_section_S) %>%
  arrange(lon)

Pac_SN <- section %>%
  filter(
    lon == parameters$lon_Pac_section,
    lat <= parameters$lat_section_N,
    lat >= parameters$lat_section_S
  ) %>%
  arrange(lat)

section_global_coordinates <- bind_rows(Atl_NS,
                     Atl_SO,
                     Pac_SO,
                     Pac_SN)

section_global_coordinates <- section_global_coordinates %>%
  mutate(lon_180 = if_else(lon > 180, lon - 360, lon))

section_global_coordinates <- section_global_coordinates %>%
  mutate(dist_int = distGeo(cbind(lon_180, lat)) / 1e6) %>%
  mutate(dist = cumsum(dist_int))


section_global_coordinates <- section_global_coordinates %>%
  select(lon, lat, dist) %>% 
  drop_na()

rm(Atl_NS, Atl_SO, Pac_SN, Pac_SO, section)

```

```{r section_global_map, fig.asp=0.6}

ggplot() +
  geom_raster(data = landmask,
              aes(lon, lat), fill = "grey80") +
  geom_point(data = section_global_coordinates,
             aes(lon, lat, col = dist)) +
  scale_fill_grey() +
  scale_colour_viridis_b(name = "Distance (Mm)") +
  coord_quickmap(expand = 0) +
  theme(legend.position = "top",
        axis.title = element_blank())

```




## Write files

```{r write_files}

basinmask_01 %>% 
  write_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                       "basin_mask_WOA18_AIP.csv"))

basinmask_01 <- basinmask_01 %>% 
  select(-basin_AIP)

basinmask_01 %>% 
  write_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                       "basin_mask_WOA18.csv"))

landmask %>% 
  write_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                       "land_mask_WOA18.csv"))

section_global_coordinates %>%
  write_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                       "section_global_coordinates.csv"))

```


# Global and regional distribution
Here we use model result for 2007 (with Normal Year Forcing, preind. atm. pCO2). 

Our variables of interest are:

- DIC
- ALK
- O2
- AOU
- NO3
- PO4
- SALT
- TEMP
- SiO3

We produced following plots for those variables:

- Map of surface value (depth = 5m, for month of Aug)
- Section- Depth plot along path from North Altantic -> Southern Ocean -> North Pacific
- Zonal mean section for each basin (Atlantic, Indian, Pacific)

## Global surface distribution

### Read netcdf
```{r read_file_and_extract_data}
# read one variable from file
nc_data <- read_ncdf(here::here("data",
                                "GIAF_JRA.pop.h.0354.nc"),
                     var = c("DIC", "ALK", "O2", "AOU", "NO3", "PO4", "SALT", "TEMP", "SiO3"))

# convert to tibble
nc_data_tibble <- nc_data %>%
  as.tibble() %>%
  mutate(nlon = if_else(nlon < 20, nlon + 360, nlon)) %>%
  mutate(z_t = z_t / 100) %>%
  rename(lon = nlon, lat = nlat, depth = z_t)

# convert as numeric
nc_data_tibble <- nc_data_tibble %>%
  mutate(DIC = as.numeric(DIC)) %>%
  mutate(ALK = as.numeric(ALK)) %>%
  mutate(O2 = as.numeric(O2)) %>%
  mutate(AOU = as.numeric(AOU)) %>%
  mutate(NO3 = as.numeric(NO3)) %>%
  mutate(PO4 = as.numeric(PO4)) %>%
  mutate(SALT = as.numeric(SALT)) %>%
  mutate(TEMP = as.numeric(TEMP)) %>%
  mutate(SiO3 = as.numeric(SiO3))

# filter Nan values
nc_data_tibble <- nc_data_tibble %>%
  filter(!is.na(DIC))

# filter surface data
nc_data_tibble_surface <- nc_data_tibble %>%
  filter(depth == 5)

# filter August data
nc_data_tibble_surface_Aug <- nc_data_tibble_surface %>%
  filter(time == as.PCICt("0355-07-30", cal = "365_day")) 

```

### Map surface DIC
```{r Map_surface_DIC}

nc_data_tibble_surface_Aug %>%
  filter(DIC > 1800) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = DIC)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "DIC Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "DIC [mmol/m3]")

```

### Map surface ALK
```{r Map_surface_ALK}

nc_data_tibble_surface_Aug %>%
  filter(ALK > 2000) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = ALK)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "ALK Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "ALK")

```

### Map surface O2
```{r Map_surface_O2}

nc_data_tibble_surface_Aug %>%
  filter(O2 > 0) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = O2)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "O2 Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "O2 [mmol/m3]")

```

### Map surface AOU
```{r Map_surface_AOU}

nc_data_tibble_surface_Aug %>%
  filter(AOU > -60) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = AOU)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "AOU Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "AOU [mmol/m3]")

```

### Map surface NO3
```{r Map_surface_NO3}

nc_data_tibble_surface_Aug %>%
  filter(NO3 > 0) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = NO3)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "NO3 Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "NO3 [mmol/m3]")

```

### Map surface PO4
```{r Map_surface_PO4}

nc_data_tibble_surface_Aug %>%
  filter(PO4 > 0) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = PO4)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "PO4 Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "PO4 [mmol/m3]")

```

### Map surface SALT
```{r Map_surface_SALT}

nc_data_tibble_surface_Aug %>%
  filter(SALT > 20) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = SALT)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "SALT Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "SALT [g/kg]")

```

### Map surface TEMP
```{r Map_surface_TEMP}

nc_data_tibble_surface_Aug %>%
  filter(TEMP > -2) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = TEMP)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "TEMP Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "TEMP [℃]")

```

### Map surface SiO3
```{r Map_surface_SiO3}

nc_data_tibble_surface_Aug %>%
  filter(SiO3 < 90) %>%
  ggplot() +
  geom_point(aes(lon, lat, col = SiO3)) +
  scale_color_viridis_c() +
  coord_quickmap(expand = 0) +
  labs(title = "SiO3 Surface Distribution", subtitle = "Depth: 5m  Month: August", x = "Longitude", y = "Latitude", colour = "SiO3 [mmol/m3]")

```

## Global section plot (U-shape)

### Define section global function
```{r define section_global function}
section_global <- function(df, var) {

  name_var <- var
  var <- sym(var)

  df_sub <- left_join(section_global_coordinates, df)

  surface <- df_sub %>%
    ggplot(aes(dist, depth, z = !!var)) +
    geom_contour_filled() +
    geom_vline(data = section_global_coordinates %>% filter(lat == 0.5),
               aes(xintercept = dist), col = "white") +
    scale_fill_viridis_d(name = name_var) +
    coord_cartesian(expand = 0,
                    ylim = c(500, 0)) +
    scale_y_reverse() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "top") +
    labs(y = "Depth (m)")

  deep <- df_sub %>%
    ggplot(aes(dist, depth, z = !!var)) +
    geom_contour_filled() +
    geom_vline(data = section_global_coordinates %>% filter(lat == 0.5),
               aes(xintercept = dist), col = "white") +
    scale_fill_viridis_d(name = name_var) +
    scale_y_reverse() +
    coord_cartesian(expand = 0, ylim = c(5400, 500)) +
    labs(x = "Distance (Mm)", y = "Depth (m)") +
    theme(legend.position = "none")

  surface / deep

}
```

### Modify data
```{r Modify_data}

rm(nc_data, nc_data_tibble_surface, nc_data_tibble_surface_Aug)
basinmask_01 <- 
  read_csv(here::here("data/World_Ocean_Atlas_2018/_summarized_files",
                       "basin_mask_WOA18_AIP.csv"))
basinmask_01 <- basinmask_01 %>%
  select(-basin)

nc_data_tibble <- nc_data_tibble %>%
  select(-time)
nc_data_tibble$lon <- round(nc_data_tibble$lon + 0.5) - 0.5
nc_data_tibble$lat <- round(nc_data_tibble$lat + 0.5) - 0.5
nc_data_tibble <- inner_join(nc_data_tibble, basinmask_01) %>%
  filter(!is.na(DIC))

```

### Global section
```{r Global_section}

section_global(nc_data_tibble, "DIC")
section_global(nc_data_tibble, "ALK")
section_global(nc_data_tibble, "O2")
section_global(nc_data_tibble, "AOU")
section_global(nc_data_tibble, "NO3")
section_global(nc_data_tibble, "PO4")
section_global(nc_data_tibble, "SALT")
section_global(nc_data_tibble, "TEMP")
section_global(nc_data_tibble, "SiO3")

```


## Zonal section plot for each basin (Atlantic, Indian, Pacific)

### Define zonal mean section function
```{r define_section_regional_function}

section_zonal <- function(df, i_basin_AIP, var) {

  name_var <- var
  var <- sym(var)

  df_sub <- df %>% 
    filter(basin_AIP == i_basin_AIP)
  
  lat_max <- max(df_sub$lat)
  lat_min <- min(df_sub$lat)

  surface <- df_sub %>%
    ggplot(aes(lat, depth, z = !!var)) +
    geom_contour_filled(breaks = breaks) +
    scale_fill_viridis_d(name = name_var) +
    coord_cartesian(expand = 0,
                    ylim = c(500, 0),
                    xlim = c(lat_min, lat_max)) +
    scale_y_reverse() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
    ) +
    labs(y = "Depth (m)",
         title = paste("Basin:", i_basin_AIP))

  deep <- df_sub %>%
    ggplot(aes(lat, depth, z = !!var)) +
    geom_contour_filled(breaks = breaks) +
    scale_fill_viridis_d(name = name_var) +
    scale_y_reverse() +
    coord_cartesian(expand = 0, 
                    ylim = c(5400, 500),
                    xlim = c(lat_min, lat_max)) +
    labs(x = "latitude (°N)", y = "Depth (m)")

  surface / deep +
    plot_layout(guides = "collect")
  
}

```

### Zonal section plot for DIC
```{r Zonal_section_DIC}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(2000, 2400, 40), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "DIC"))
  
}

```

### Zonal section plot for ALK
```{r Zonal_section_ALK}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(2000, 2600, 50), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "ALK"))
  
}

```

### Zonal section plot for O2
```{r Zonal_section_O2}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(0, 500, 50), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "O2"))
  
}

```

### Zonal section plot for AOU
```{r Zonal_section_AOU}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(-50, 350, 40), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "AOU"))
  
}

```

### Zonal section plot for NO3
```{r Zonal_section_NO3}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(0, 45, 5), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "NO3"))
  
}

```

### Zonal section plot for PO4
```{r Zonal_section_PO4}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(0, 3.2, 0.4), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "PO4"))
  
}

```

### Zonal section plot for SALT
```{r Zonal_section_SALT}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(20, 40, 2), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "SALT"))
  
}

```

### Zonal section plot for TEMP
```{r Zonal_section_TEMP}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(-3, 33, 4), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "TEMP"))
  
}

```

### Zonal section plot for SiO3
```{r Zonal_section_SiO3}

for (i_basin_AIP in unique(nc_data_tibble$basin_AIP)) {
    
    breaks <- c(seq(-20, 90, 10), Inf)
    print(section_zonal(nc_data_tibble,
                        i_basin_AIP = i_basin_AIP,
                        "SiO3"))
  
}

```
